---
title: "Autocorrelation and Normality Tests"
subtitle: "Empirical Distribution of S&P500 Returns"
author: "Shirley He"
output:
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---


# Notebook Contents   
  - [Clean up, Load Libraries](#load_libraries)  
  - [ADF on &$P500 Price](#adf_prices)   
  - [ADF on S&P500 Returns](#adf_returns)  
  - [Histogram, QQ Plot, Kernel Histogram](#hist_qqplot_kernel)  
  - [Descriptive Statistics and JB Test](#descriptive_stats_jb)   
  - [ACF, PACF, and Ljung-Box Test](#acf_pacf)   
  - [Finish Up with Virtual Environment](#finish)

# Clean up and load libraries{#load_libraries}
```{r}
rm( list=ls() )
library(tseries)
library(moments)
```

# ADF Test on Price Series{#adf_prices}
The p-value from the Augmented Dickey-Fuller test is 0.9249  
The null of ADF test is that the data are non-stationary.   
We want to reject the null hypothesis for this test, so we want a p-value of 0.05 or smaller.  

We hope that if we take log differences of this, the p-value will go down

 

```{r}
D <- read.table('../data/sp500.csv',header=TRUE,sep=',')
colnames(D)
#reverse adjusted close column. take it as sp500
sp500 <- rev(D$adjusted_close)

adf.test(sp500) #p-value = 0.9249

```





# ADF Test on Returns Series{#adf_returns}
The p-value from the Augmented Dickey-Fuller test is 0.01  


```{r}
D_returns <- read.table('../data/sp500-returns.csv');

#2 columns: dates and returns
dates <- as.Date(as.character(D_returns[,1]),'%Y-%m-%d')
ret   <- D_returns[,2]

adf.test(ret)
```

 
 
# Histogram, QQ Plot and Kernel Histogram of Returns{#hist_qqplot_kernel}

### Histogram
When we have time series data, the first thing to do is plot a histogram.  
We see that the data is clearly not gaussian.  

Gaussian has some observations at 5, -5, 10, -10 so the tails are not exponentially decaying like in normal.  
If this were gaussian, the likelihood of seeing 5 should be almost never.  
Even if we play with the mean and variance (increase variance to flatten it), we could still not fit a gaussian in here. 


```{r}
hist(ret,50,freq=FALSE,col='tomato',border='darkred',main='') #main=title
```

## QQ Plot
Ifthe data is Gaussian, we know exactly the numbers of the quantiles   
Here we compute the sample quantiles from our data and compare it to gaussian quantiles   
```{r}
qqnorm(ret,col='tomato',main='')
qqline(ret,lwd=2,lty=3)
```

## Kernel Histogram
We need an empirical estimate of the distribution.  

Notes: 
  - 'density' is the kernel function in R   
  - 'polygon' function needs a start and end to fill, but passing kernel to it will fill what's in between the area under the shape 

```{r}
kernel <- density(ret) 
kernel

#kernel object has a plot method. so when you call plot on it, it will plot something already for you. 
plot(kernel , main='' , yaxs='i' ) 
polygon(kernel , col="tomato" , border='darkred' ) #we fill it using polygon. 

lines( seq(-10,10,0.1) , dnorm( seq(-10,10,0.1) , mean(ret) , sd(ret) ) , col='darkblue' ,lwd=4) #compare to density of normal with mean 10 (we know how that normal should look)

```

 
# Descriptive Stats and JB Test{#descriptive_stats_jb}
Create vector of mean returns, annualised volatility, skewness and kurtosis

```{r}
descr <- as.matrix( c( mean(ret) , sqrt(252)*sd(ret) , skewness(ret) , kurtosis(ret) ) ) 
dimnames(descr)[[1]] <- list('mean','ann vol','skew','kurt')
round(descr, 3)
```

## JB Test
The JB Test compares the density to a standard normal.   
The p-value from this test is < 2.2e-16 (very small) which tells us that the data is not normal.   

Kurtosis is almost 3x the standard normal, which is 3. 

```{r}
jarque.test(ret)

```

# ACF, PACF, and Ljung-Box Test{#acf_pacf}
## ACF 
Auto-correlation function finds correlations between the present and lags
The ACF graph shows that there is not much autocorrelation

```{r}

ret.acf <- acf(ret , ylim=c(-0.1,1) , lwd=5 , xlim=c(0,25) , col='darkorange2')
round(head(ret.acf$acf, 5), 3) #going into ret.acf function, tell it what to look for.
ret.acf

```

## PACF
Partial auto-correlation function finds correlation of the residuals which can remain after removing the effects which are already explained by other lags. 

```{r}
#compared to ACF, here we partial out all past influences up to k (??)
ret.pacf <- pacf(ret, ylim=c(-0.1,1) , lwd=5 , xlim=c(0,25) , col='darkorange2')
ret.pacf
```

## Ljung-Box Test
A significant p-value in this test rejects the null hypothesis that the time series isn’t autocorrelated  

```{r}
Box.test( ret , lag=22 , type="Ljung-Box" )
```

# Finish Up with Virtual Environment renv{#finish}
Save state of project library to `renv.lock`
```{r}
renv::snapshot()
```
