---
title: "The Rejection Region when Data is Autocorrelated"
subtitle: "The Z-test for Time Series"
author: "Shirley He"
output:
  html_document:
    toc: true
    toc_float: 
      toc_collapsed: true
    toc_depth: 3
    number_sections: true
    theme: lumen
---


# Notebook Contents 
 - [Setup Montecarlo](#montecarlo)
 - [Monte Carlo 1 - iid Data](#iid_data)
 - [Monte Carlo 2 - Dependent data](#dependent_data)
 
 
# Setup Montecarlo{#montecarlo} 
```{r}
rm(list=ls())
set.seed(42)
S <- 10000
T <- 500
crit <- qnorm(0.95)
w   <- c(0.4,0.3,0.2)
```

## Carry one replication of the exeperiment
```{r}
eps <- rnorm(T)
y   <- rep(0,T)

for( t in 4:T ){
  y[t] <- eps[t] + w[1]*eps[t-1] + w[2]*eps[t-2] + w[3]*eps[t-3]
}

m      <- mean(y)
s2     <- var(y)
se.iid <- sqrt(s2/T)

gamma0 <- var(y)
gamma1 <- cov(y[2:T],y[1:(T-1)])
gamma2 <- cov(y[3:T],y[1:(T-2)])
gamma3 <- cov(y[4:T],y[1:(T-3)])
s2.lr  <- gamma0 + 2*(gamma1 + gamma2 + gamma3)
se.ts  <- sqrt(s2.lr/T)

t1 <- m/se.iid
t2 <- m/se.ts

par( mar=c(2,2,0.75,0.75) , yaxs='i' )
plot( seq(-4,4,0.1) , dnorm(seq(-4,4,0.1)) , col='darkblue' , lwd=3 , t='l',ylim=c(0,0.5),xlim=c(-8,8) )
polygon( c(crit,crit,10,10) , c(0,0.5,0.5,0) ,col='coral2' , density=10)
abline( v=qnorm(0.95) , col='firebrick3' , lwd=4)
text(crit,0.25,sprintf('Nominal 5%% Rejection Region'),col='black',pos=4)
box()
```

# Monte Carlo - 1 iid Data{#iid_data}

Check out the distibution of the test statistic for iid data when the data has serial dependence
```{r}
z.score <- rep(0,S)

for( s in 1:S ){
	eps <- rnorm(T)
	y   <- rep(0,T)

	for( t in 4:T ){
		y[t] <- eps[t] + w[1]*eps[t-1] + w[2]*eps[t-2] + w[3]*eps[t-3]
	}
  
  m  <- mean(y)
  s2 <- var(y)
  se <- sqrt(s2/T)
  
  z.score[s] <- m/se
}

```

`z.size` is around 16%
This means that the rejection region is around 16% instead of 5%  
```{r}
z.size <- mean( z.score > crit ) 
```


Plot the rejection region over the histogram
```{r}
par( mar=c(2,2,0.75,0.75) , yaxs='i' )
hist(z.score,50,freq=FALSE,ylim=c(0,0.5),xlim=c(-8,8),col='azure2',main='')
lines( seq(-4,4,0.1) , dnorm(seq(-4,4,0.1)) , col='darkblue' , lwd=3 , t='l',ylim=c(0,0.5),xlim=c(-8,8) )
polygon( c(crit,crit,10,10) , c(0,0.5,0.5,0) ,col='coral2' , density=10)
abline( v=qnorm(0.95) , col='firebrick3' , lwd=4)
text(qnorm(0.95),0.25,sprintf('Nominal 5%% Rejection Region:\nSize %2.2f%%',z.size*100),col='black',pos=4)
box()

```

# Monte Carlo 2 - Dependent data{#dependent_data}
Check out the distibution of the proposed test statistic for dependent data when the data has serial dependence
```{r}
z.score <- rep(0,S)

for( s in 1:S ){
	eps <- rnorm(T)
	y   <- rep(0,T)

	for( t in 4:T ){
		y[t] <- eps[t] + w[1]*eps[t-1] + w[2]*eps[t-2] + w[3]*eps[t-3]
	}
  
  m  <- mean(y)
  gamma0 <- var(y)
  gamma1 <- cov(y[2:T],y[1:(T-1)])
  gamma2 <- cov(y[3:T],y[1:(T-2)])
  gamma3 <- cov(y[4:T],y[1:(T-3)])
  s2.lr  <- gamma0 + 2*(gamma1 + gamma2 + gamma3)

  se <- sqrt(s2.lr/T)
  
  z.score[s] <- m/se

}
```


`z.size` is around 5%  
This is much closer to the nominal rejection region
```{r}
z.size <- mean( z.score > crit )

par( mar=c(2,2,0.75,0.75) , yaxs='i' )
hist(z.score,50,freq=FALSE,ylim=c(0,0.5),xlim=c(-8,8),col='azure2',main='')
lines( seq(-4,4,0.1) , dnorm(seq(-4,4,0.1)) , col='darkblue' , lwd=3 , t='l',ylim=c(0,0.5),xlim=c(-8,8) )
polygon( c(crit,crit,10,10) , c(0,0.5,0.5,0) ,col='coral2' , density=10)
abline( v=qnorm(0.95) , col='firebrick3' , lwd=4)
text(qnorm(0.95),0.25,sprintf('Nominal 5%% Rejection Region:\nSize %2.2f%%',z.size*100),col='black',pos=4)
box()
```
